---
title: "The Story of Sydney"
subtitle: "A 146-Year Demographic Journey & Future Projections of a Classic Name"
author: "Sydney Martin"
date: today
date-format: "MMMM D, YYYY"
format:
  html:
    theme: zephyr
    toc: true
    toc-depth: 2
    toc-title: "Overview"
    smooth-scroll: true
    code-fold: true
    code-summary: "View R Code"
execute:
  warning: false
  message: false
---

```{=html}
<style>
@import url('https://fonts.googleapis.com/css2?family=Plus+Jakarta+Sans:wght@300;400;500;600;700;800&family=Playfair+Display:ital,wght@0,600;0,700;1,400&display=swap');

body {
  font-family: 'Plus Jakarta Sans', -apple-system, BlinkMacSystemFont, sans-serif;
  color: #2d3748;
  background-color: #fafbfc;
  line-height: 1.7;
}

header#title-block-header {
  margin-top: 1.5rem;
  margin-bottom: 2rem;
  padding-bottom: 1.5rem;
  border-bottom: 1px solid #edf2f7;
}

h1.title {
  font-family: 'Playfair Display', Georgia, serif;
  font-weight: 700;
  font-size: 2.75rem;
  color: #1e293b;
  letter-spacing: -0.02em;
}

p.subtitle {
  font-size: 1.15rem;
  color: #64748b;
  font-weight: 400;
  margin-top: 0.5rem;
}

.quarto-title-meta {
  color: #94a3b8;
  font-size: 0.9rem;
}

h2 {
  font-weight: 700;
  color: #1e293b;
  margin-top: 2.25rem;
  padding-bottom: 0.5rem;
  border-bottom: 1px solid #f1f5f9;
}

/* Stat summary cards */
.stat-grid {
  display: grid;
  grid-template-columns: repeat(auto-fit, minmax(160px, 1fr));
  gap: 1rem;
  margin: 1.75rem 0 2rem 0;
}

.stat-box {
  background: white;
  padding: 1.25rem 1rem;
  border-radius: 12px;
  box-shadow: 0 4px 16px rgba(0, 0, 0, 0.03);
  border: 1px solid #e2e8f0;
  text-align: center;
  transition: transform 0.2s ease, box-shadow 0.2s ease;
}

.stat-box:hover {
  transform: translateY(-2px);
  box-shadow: 0 8px 24px rgba(0, 0, 0, 0.06);
}

.stat-value {
  font-size: 1.85rem;
  font-weight: 800;
  color: #6366f1;
  display: block;
  letter-spacing: -0.02em;
}

.stat-label {
  font-size: 0.78rem;
  color: #64748b;
  text-transform: uppercase;
  letter-spacing: 0.06em;
  font-weight: 600;
  margin-top: 0.35rem;
}

/* Card wrapper for plots */
.plot-card {
  background: white;
  padding: 1rem;
  border-radius: 14px;
  box-shadow: 0 4px 20px rgba(0, 0, 0, 0.03);
  border: 1px solid #e2e8f0;
  margin: 1.25rem 0 1.5rem 0;
}

.plotly.html-widget {
  border-radius: 8px;
}
</style>
```

```{=html}
<div class="stat-grid">
  <div class="stat-box">
    <span class="stat-value">10,243</span>
    <span class="stat-label">Peak Births (2000)</span>
  </div>
  <div class="stat-box">
    <span class="stat-value">0.56%</span>
    <span class="stat-label">Peak Share of Girls</span>
  </div>
  <div class="stat-box">
    <span class="stat-value">873</span>
    <span class="stat-label">Births in 2025</span>
  </div>
  <div class="stat-box">
    <span class="stat-value">~350</span>
    <span class="stat-label">2035 Projection</span>
  </div>
</div>
```

## Setup & Data Pipeline

The data is sourced directly from the **US Social Security Administration**, containing records of every registered birth in the United States from 1880 through 2025.

```{r}
#| label: setup
library(tidyverse)
library(plotly)

# List all annual text files in the data directory
files <- fs::dir_ls("data", glob = "*yob*.txt")

# Read, aggregate, and calculate annual proportions among female births
babies <- read_csv(
  files,
  id = "file",
  col_names = c("name", "sex", "count"),
  show_col_types = FALSE
) |>
  mutate(year = parse_number(file)) |>
  select(-file) |>
  filter(sex == "F") |>
  group_by(year) |>
  mutate(prop = count / sum(count)) |>
  ungroup()
```

## The Last Decade (2016–2025)

Over the past ten years, the popularity of the name **Sydney** has experienced a steady, noticeable decline. 

::: {.callout-note appearance="minimal"}
💡 **Interactive Chart:** Hover or click anywhere on the line to inspect the exact number of babies and the percentage share for each year.
:::

```{=html}
<div class="plot-card">
```

```{r}
#| label: sydney-last-10-years
sydney_recent <- babies |>
  filter(name == "Sydney", year >= 2016) |>
  mutate(tooltip = paste0(
    "<b>Year:</b> ", year, "<br>",
    "<b>Count:</b> ", scales::comma(count), " babies<br>",
    "<b>Share:</b> ", scales::percent(prop, accuracy = 0.01)
  ))

p_recent <- ggplot(sydney_recent, aes(x = year, y = prop, text = tooltip, group = 1)) +
  geom_line(color = "#6366f1", linewidth = 1.3) +
  geom_point(color = "#4f46e5", size = 2.8) +
  scale_x_continuous(breaks = 2016:2025) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.01)) +
  labs(
    title = 'Popularity of Female Name "Sydney" (2016–2025)',
    x = "Year",
    y = "Percentage of Female Births"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 13, color = "#1e293b"),
    axis.title = element_text(size = 11, color = "#64748b"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "#f1f5f9")
  )

ggplotly(p_recent, tooltip = "text") |>
  config(displayModeBar = FALSE)
```

```{=html}
</div>
```

### Key Takeaways:
- **Halved in 10 Years:** In 2016, Sydney represented **0.13%** of all female births (2,322 babies). By 2025, that proportion fell to **0.05%** (873 babies).
- **Consistent Trajectory:** Aside from a slight uptick in 2024, the downward trend has been remarkably steady year-over-year.

---

## 146 Years in Perspective (1880–2025)

Placing the last decade into historical context reveals an incredible story of cultural emergence, explosive popularity, and subsequent settling.

```{=html}
<div class="plot-card">
```

```{r}
#| label: sydney-full-history
sydney_history <- babies |>
  filter(name == "Sydney") |>
  mutate(tooltip = paste0(
    "<b>Year:</b> ", year, "<br>",
    "<b>Count:</b> ", scales::comma(count), " babies<br>",
    "<b>Share:</b> ", scales::percent(prop, accuracy = 0.01)
  ))

p_history <- ggplot(sydney_history, aes(x = year, y = prop, text = tooltip, group = 1)) +
  geom_area(fill = "#e0e7ff", alpha = 0.5) +
  geom_line(color = "#6366f1", linewidth = 1.3) +
  scale_x_continuous(breaks = seq(1880, 2025, by = 20)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.1)) +
  labs(
    title = 'Historical Trajectory of Female Name "Sydney" (1880–2025)',
    x = "Year",
    y = "Percentage of Female Births"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 13, color = "#1e293b"),
    axis.title = element_text(size = 11, color = "#64748b"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "#f1f5f9")
  )

ggplotly(p_history, tooltip = "text") |>
  config(displayModeBar = FALSE)
```

```{=html}
</div>
```

### Two Defining Patterns:
1. **The Turn-of-the-Century Phenomenon:** Sydney was virtually unheard of as a female name for most of the 1900s (under 0.01%). Starting in the late 1980s, it surged dramatically, peaking in **2000** at **0.56%** of all female births (over 10,200 babies).
2. **The Modern Correction:** Since its peak around 2000, the name has seen an orderly decline, moving toward a low and stable long-term equilibrium.

---

## Future Projections (2026–2035)

Where is the name headed next? We modeled the recent ten-year trend using an **exponential decay model**, which captures natural stabilization as popularity approaches baseline levels.

```{=html}
<div class="plot-card">
```

```{r}
#| label: sydney-projection
# Fit exponential model on 2016-2025 observations
fit_exp <- lm(log(prop) ~ year, data = sydney_recent)

# Baseline annual births for count estimations
birth_baseline <- 1607000

# 10-year projection
future_years <- tibble(year = 2026:2035) |>
  mutate(
    prop = exp(predict(fit_exp, newdata = tibble(year = 2026:2035))),
    count = round(prop * birth_baseline),
    type = "Projected",
    tooltip = paste0(
      "<b>Year:</b> ", year, " (Projected)<br>",
      "<b>Est. Count:</b> ~", scales::comma(count), " babies<br>",
      "<b>Share:</b> ", scales::percent(prop, accuracy = 0.01)
    )
  )

historical_recent <- sydney_recent |>
  select(year, prop, count, tooltip) |>
  mutate(type = "Observed")

combined_trend <- bind_rows(historical_recent, future_years)

p_proj <- ggplot(combined_trend, aes(x = year, y = prop, color = type, text = tooltip, group = type)) +
  geom_line(linewidth = 1.3) +
  geom_point(size = 2.8) +
  scale_color_manual(values = c("Observed" = "#6366f1", "Projected" = "#f43f5e")) +
  scale_x_continuous(breaks = seq(2016, 2035, by = 2)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.01)) +
  labs(
    title = '10-Year Forecast for "Sydney" (2026–2035)',
    x = "Year",
    y = "Percentage of Female Births",
    color = "Data Series"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 13, color = "#1e293b"),
    axis.title = element_text(size = 11, color = "#64748b"),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "#f1f5f9")
  )

ggplotly(p_proj, tooltip = "text") |>
  config(displayModeBar = FALSE)
```

```{=html}
</div>
```

### Forecast Insights:
- **Projected Frequency:** By **2035**, Sydney is anticipated to represent around **0.02%** of newborn girls (~350 babies per year in the US).
- **Decay vs. Linear:** A naive linear model would predict impossible negative numbers by 2032. The exponential model accurately reflects how name usage tapers gradually into a steady classic niche.
