---
title: "Popularity of the Name Sydney"
author: "Sydney Martin"
format:
  html:
    theme: flatly
    code-fold: true
    toc: true
execute:
  warning: false
  message: false
---

## Setup and Data Loading

```{r}
#| label: setup
library(tidyverse)
library(plotly)

# Find all annual birth files
files <- fs::dir_ls("data", glob = "*yob*.txt")

# Read in each year of data and calculate annual female proportions
babies <- read_csv(
  files,
  id = "file",
  col_names = c("name", "sex", "count"),
  show_col_types = FALSE
) |>
  mutate(year = parse_number(file)) |>
  select(-file) |>
  filter(sex == "F") |>
  group_by(year) |>
  mutate(prop = count / sum(count)) |>
  ungroup()
```

## Trend Over the Last 10 Years (2016–2025)

Hover or click on any point to see the exact count and percentage for that year.

```{r}
#| label: sydney-last-10-years
sydney_recent <- babies |>
  filter(name == "Sydney", year >= 2016) |>
  mutate(tooltip = paste0(
    "<b>Year:</b> ", year, "<br>",
    "<b>Count:</b> ", scales::comma(count), " babies<br>",
    "<b>Share:</b> ", scales::percent(prop, accuracy = 0.01)
  ))

p_recent <- ggplot(sydney_recent, aes(x = year, y = prop, text = tooltip, group = 1)) +
  geom_line(color = "#2c7bb6", linewidth = 1.2) +
  geom_point(color = "#2c7bb6", size = 2.5) +
  scale_x_continuous(breaks = 2016:2025) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.01)) +
  labs(
    title = 'Popularity of Female Name "Sydney" (2016–2025)',
    x = "Year",
    y = "Percentage of Female Births"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

ggplotly(p_recent, tooltip = "text")
```

Over the last 10 years, the proportion of baby girls named **Sydney** has steadily fallen by more than half:
- In **2016**, **0.13%** of all female births were named Sydney (2,322 babies).
- By **2025**, that figure dropped to **0.05%** (873 babies).

## Historical Patterns (1880–2025)

Hover over the line to explore counts and percentages throughout history.

```{r}
#| label: sydney-full-history
sydney_history <- babies |>
  filter(name == "Sydney") |>
  mutate(tooltip = paste0(
    "<b>Year:</b> ", year, "<br>",
    "<b>Count:</b> ", scales::comma(count), " babies<br>",
    "<b>Share:</b> ", scales::percent(prop, accuracy = 0.01)
  ))

p_history <- ggplot(sydney_history, aes(x = year, y = prop, text = tooltip, group = 1)) +
  geom_line(color = "#2c7bb6", linewidth = 1.2) +
  scale_x_continuous(breaks = seq(1880, 2025, by = 20)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.1)) +
  labs(
    title = 'Historical Popularity of Female Name "Sydney" (1880–2025)',
    x = "Year",
    y = "Percentage of Female Births"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

ggplotly(p_history, tooltip = "text")
```

Looking at the entire 146-year dataset, two striking patterns stand out:

1. **The 1990s Boom & 2000 Peak:** For most of the 20th century, Sydney was extremely rare as a female name (representing under **0.01%** of baby girls). Starting around 1990, its popularity surged dramatically, reaching an all-time peak in **2000** at **0.56%** of all baby girls (10,243 births).
2. **Modern Decline:** Following that 2000 summit, the name has seen a steady, continuous downward trend each decade, settling at about one-tenth of its peak level today.

## Future Projections (2026–2035)

Hover over the observed vs. projected lines to inspect specific values.

```{r}
#| label: sydney-projection
# Fit an exponential decay model based on the 2016-2025 trajectory
fit_exp <- lm(log(prop) ~ year, data = sydney_recent)

# Estimated annual baseline of female births (~1.61M) to project baby counts
birth_baseline <- 1607000

# Create a projection for the next 10 years
future_years <- tibble(year = 2026:2035) |>
  mutate(
    prop = exp(predict(fit_exp, newdata = tibble(year = 2026:2035))),
    count = round(prop * birth_baseline),
    type = "Projected",
    tooltip = paste0(
      "<b>Year:</b> ", year, " (Projected)<br>",
      "<b>Est. Count:</b> ~", scales::comma(count), " babies<br>",
      "<b>Share:</b> ", scales::percent(prop, accuracy = 0.01)
    )
  )

# Combine observed and projected data
historical_recent <- sydney_recent |>
  select(year, prop, count, tooltip) |>
  mutate(type = "Observed")

combined_trend <- bind_rows(historical_recent, future_years)

p_proj <- ggplot(combined_trend, aes(x = year, y = prop, color = type, text = tooltip, group = type)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.5) +
  scale_color_manual(values = c("Observed" = "#2c7bb6", "Projected" = "#d7191c")) +
  scale_x_continuous(breaks = seq(2016, 2035, by = 2)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.01)) +
  labs(
    title = 'Projected Trend for "Sydney" (2026–2035)',
    x = "Year",
    y = "Percentage of Female Births",
    color = "Series"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

ggplotly(p_proj, tooltip = "text")
```

Based on an exponential decay model fitted to recent 2016–2025 data:
- **Projected Share:** The share of baby girls named Sydney is projected to gradually drop from **0.05%** in 2025 to approximately **0.02%** by 2035 (around 350 babies per year).
- **Why Exponential Decay?** A simple linear model would unrealistically predict negative percentages after 2031. The exponential model captures the natural slowing rate of decline as the name approaches a low, stable baseline.
